---
title: "Misclassification Activity for JSM 2023 Short Course"
author: "set below"
date: "set below"
output: 
  beamer_presentation:
    theme: "Boadilla"
    colortheme: "orchid"
header-includes:
  - \AtBeginDocument{\title[]{Misclassification Activity\\JSM 2023 Short Course}}
  - \AtBeginDocument{\author[]{Paul Gustafson}}
  - \AtBeginDocument{\date[]{August 6, 2023}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=F}
### global reproducibility
set.seed(1234)
### sane output
options(digits=3)
options(width=65)
```

## Recall this example (slide 128)

```{r}
data.tbl <- matrix(c(45,94,257,945),
  dimnames = list(c("CHD+", "CHD-"),c("Resin+", "Resin-")),
  nrow = 2, byrow = TRUE)
```

```{r}
data.tbl
```
## Naive analysis presuming correct exposure classification

Inference for exposure-disease odds-ratio

```{r}
logOR.hat <- sum(c(1,-1,-1,1)*log(as.vector(data.tbl)))

logOR.SE <- sqrt(sum(1/as.vector(data.tbl)))

exp(logOR.hat + c(0, -1.96, 1.96)*logOR.SE)
```

## Assuming nondifferential exposure misclassification with 90% sensitivity and 80% specificity

Recall from slides 138, 139:

```{r, message=F, warning=F}
require(episensr)

ft <- misclassification(data.tbl,
 type="exposure", bias_parms=c(0.8, 0.8, 0.9, 0.9))

# point and interval estimation of OR
ft$adj.measures[2,]
```


## Activity A 

Check you can reproduce one of the **differential** classification adjustments on slide 135 (i.e., one of the off-diagonal table entries).

For instance, try presuming 90% specificity for all subjects, but sensitivity of 90% for controls, compared to 80% for cases.

Might help:

```{r,eval=F}
help(misclassification)
```



## Activity B: Uncertainty about misclassification rates. 

Say the investigator is confident that the misclassification is nondifferential.    

Has 85% sensitivity and 85% specificity as "best guesses."  

But thinks either guess could be off by as much as five percentage points.

Can you look at
```{r, eval=F}
help(probsens)
```
And provide an appropriate analysis?

HINT: First example in the help gives a template.

HINT: For simplicity, maybe "triangular" or "uniform" instead of "trapezoidal"





## Activity C - Role of data

We have useful heuristics in statistics, such as the primal role of $\sqrt{n}$.    

If I want interval estimates *twice* as narrow, I likely need about *four times* as much data.

Repeat Activity B, but with four times as much data.   (Simplest to just keep cell *proportions* fixed in the 2 by 2 data table).

Reflect on what you find.


 




